---
title: "README"
author: "Marieke Dirksen"
date: "October 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RSOI
## Reduced space optimal interpolation workshop 2019

For an interpolation area the following files need to be prepared:

* empty grid layers for each month of the reconstruction period
* observations for the 30 year reference period
* observations for the reconstruction period

## An overview of the observations

### Indonesia

```{r}
library(mapview)
data("sp.RR_stations")
mapview(sp.RR_stations)
```

### The Netherlands

```{r}
data("sp.TN_stations")
mapview(sp.TN_stations)

```

## Datasets in this project
This project contains meteorological data from two countries: the Netherlands and Indonesia. From the Netherlands temperature observations since 1905 are included. For Indonesia precipitation observations are stored in `data("RR_obs")` starting in the year 1930. 

```{r}
#load Temperature data for the Netherlands
data("TN_observations") #observations
data("TN_reference")    #reference observations
data("sp_grid_nl")      #empty array
#data("ok_NL")           #ordinary kriging climatology (not used in RSOI)

#load Precipitation data for Indonesia
data("RR_obs")          #observations
data("RR_ref")          #reference observations
data("sp_grid_ind")     #empty array
#data("RR_ref_grid")     #reference grid interpolated (not used in RSOI)
```
